Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-26 paper/paper.Rmd
close #7 by adding in changes and editting

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts and being trusted by policy-makers.

At the same time, ensemble forecasts have a range of potential downsides.
The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
Each of these may impact the long-term quality of the resulting ensemble forecasts.

Here, we focus on addressing the issues of resource cost, and interpretability.
In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@Iskauskas2022-fz]).

In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources with an easily generalised implementation.

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the US forecasting hub [@Cramer2022-sp] in early 2020, the Germany/Poland forecast hub [@Bracher2021-wk] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@Sherratt2022-ue], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF; @Bosse2022-hm].
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@Bosse2022-hm].
We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts [@Bosse2022-hm].

Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm]. 
Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts. They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence. On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts. Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting. There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.

Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance. 
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.

In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions. 
We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble. 
First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.  
We briefly outline some of the summary statistics it produces and their epidemiological relevance.  We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.  
Finally, we summarise our findings, discuss their implications, and highlight areas for potential study. We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.

>>>>

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakholders.

Whilst there is robust and consistent evidence that ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
Each of these may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality. Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study). All of these collaborations, which the author have been involved with managing, ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate.

In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justified assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a large scale forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz] utility, it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.

In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a less degree targets.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts. They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence. On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts. Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting. There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models

Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.

@@

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model, where variance is linked to the mean, is a sensible choic. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].

In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions.
We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.  We briefly outline some of the summary statistics it produces and their epidemiological relevance.
We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.  Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
